---
title: "Green Taxi Case Study"
author: "Chengran Ouyang"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: lumen
    number_sections: true
    font-family: Open Sans, sans-serif
    font-import: https://fonts.googleapis.com/css?family=Open+Sans
    code_folding: hide
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Define the Problem
**Requirment:**

1. Download the data, load it into your favorite statistical programing software or database.
Report the number of rows and columns that you've loaded.

2. Visualize trip distance by time of day in any way you see fit, any observations?

3. What are the most popular pickup locations on weekdays vs weekend?

4. I want to know where I can most easily get a cab. Recommend a pick up spot where i can
find a cab given my lat long.

This time, I would use **R** as the programing software for the analysis. The analysis would be based on a standard data science framework and the questions above; however, I would extend the scope of the analysis to identify any unique insight as well as provide detailed explaination of my code.

The Process of Coding is available via [my github account](https://github.com/couyang24?tab=repositories).

# Step 2: Gather the Data
The scope of data is limited to Green taxi data for February 2016. There is also data dictionary available from the website, which explain the variable. The Data is available via [NYC Trip Data](http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml).

# Step 3: Preprocess the Data

## Dependencies {.tabset}
### Required libraries

```{r }
# pacman package install required packages if not available 
if (!require("pacman")) install.packages("pacman") 
pacman::p_load(tidyverse, DT, lubridate, leaflet, leaflet.extras, maps, data.table, ggthemes, rebus)
```

### Required Dataset

```{r }
# Initially use read.csv then write the file so that going forward I can use fread
# data <- read.csv("green_tripdata_2016-02.csv", stringsAsFactors = F)
# write_csv(data,"green_tripdata_2016-02v2.csv")
data <- fread("green_tripdata_2016-02v2.csv", stringsAsFactors = F, data.table = FALSE, na.strings=c("NA","NaN","?", ""),
              showProgress = F)
```

## First Glimpse {.tabset}
### First 20 rows with selected columns

```{r }
data %>% head(20) %>% select(lpep_pickup_datetime,Pickup_longitude,Pickup_latitude,Trip_distance) %>% datatable()
```

### Structure


The second question can be answered by looking at the structure of the dataset. The dataset has `r nrow(data)` observations(rows) 
and `r ncol(data)` variables(columns).


```{r }
data %>% str() # 1510722 obs. of  21 variables
```

### Summary

```{r }
data %>% summary()
```

## Data Cleaning: Correcting, Completing, Creating, and Converting {.tabset}
### Correcting & Completing


Since the trip payment is not in the scole of this analysis, I took out these variables for shorter runing time.


```{r }
data[,which(str_detect(names(data),"amount|fee|Extra|fee|Pay|tax|charge"))] <- NULL
```


Looking at the summary result, most value in pickup_longitude is around -73.95 (Presumbly where New York is) but the
variable also have 0 and -115.28. Then, I put the pickup location on the map below, we see the locations in Guinea Basin
and Las Vegas. The Guinea Basin location is due to having 0 longtitude and latitude. The Las Vegas locations seem interesting
for which, might worth some time to dig into but it is beyond the scope of this analysis.


```{r }
set.seed(0)
data %>% sample_n(size=10000) %>% 
  leaflet() %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addCircleMarkers(~Pickup_longitude, ~Pickup_latitude, radius = 1,
                   color = "firebrick", fillOpacity = 0.001)
```



Therefore, I limited the boundary to New York and got rid out these locations. Then, I got the map below. It is very interesting
to see that all the pick up location are outside of the core area of new york city. By doing a little research, I found out that
the green taxi are only allowed to pick up passengers (street hails or calls) in outer boroughs (excluding John F. Kennedy
International Airport and LaGuardia Airport unless arranged in advance) and in Manhattan above East 96th and West 110th Streets.
That explains the pattern we see here.


```{r }
data <- data %>% filter(Pickup_longitude > -75, Pickup_longitude < -73, Pickup_latitude > 40.4, Pickup_latitude <41,
                        Dropoff_longitude!=0,Dropoff_latitude!=0)

set.seed(0)
data %>% sample_n(size=10000) %>% 
  leaflet() %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addCircleMarkers(~Pickup_longitude, ~Pickup_latitude, radius = 1,
                   color = "firebrick", fillOpacity = 0.001)

```

### Creating, and Converting


I converted pick_datetime to time data and created variables such as pickup_hour, pickup_weekday, pickup_weekend, etc. 


pickup_hour has value from 1 to 24, denoting 24 hours a day.


pickup_weekday has value from Monday to Friday and is categorized as factor.


pickup_weekend has value Weekday and Weekend.


```{r }
data <- data %>% mutate(lpep_pickup_datetime = ymd_hms(lpep_pickup_datetime),
                pickup_hour=hour(lpep_pickup_datetime)+1,
                pickup_weekday=as.factor(weekdays(lpep_pickup_datetime)),
                pickup_weekend=if_else(pickup_weekday=='Saturday'|pickup_weekday=='Sunday','Weekend','Weekday'),
                Lpep_dropoff_datetime = ymd_hms( Lpep_dropoff_datetime),
                dropoff_hour=hour(lpep_pickup_datetime)+1,
                dropoff_weekday=as.factor(weekdays(lpep_pickup_datetime)))
```

# Step 4 Perform Exploratory Data Analysis (EDA)

## The Distribution of Distance {.tabset}

### Raw

First of all, let's have a overview on the distance data. As the graph below, the data is skewed by the distance outliers. 


```{r}
data %>%
  ggplot(aes(Trip_distance)) +
  geom_histogram(fill = "firebrick", bins = 150) +
  theme_economist() +
  labs(title='Green Taxi Case Study',subtitle='Initial Historgram of Distance',caption="source: NYC Green Taxi Data",
       x="Travel Distance")
```


### Processed


After processing the data, I got an almost normal distribution with some more few distance trips.


```{r}
data %>%
  ggplot(aes(Trip_distance)) +
  geom_histogram(fill = "firebrick", bins = 150) +
  scale_x_log10() +
  theme_economist() +
  labs(title='Green Taxi Case Study',subtitle='Processed Historgram of Distance',caption="source: NYC Green Taxi Data",
       x="Travel Distance (log)")
```


## Visualize Trip Distance by Time of Day





## Top 5 Pick Up Locations on Weekdays and Weekend


```{r }
round_num <- 3

Weekend_Top5 <- data %>% filter(pickup_weekend=='Weekend') %>% 
  group_by(lng=round(Pickup_longitude,round_num),lat=round(Pickup_latitude,round_num)) %>% 
  count() %>% arrange(desc(n)) %>% head(5)

Weekday_Top5 <- data %>% filter(pickup_weekend=='Weekday') %>% 
  group_by(lng=round(Pickup_longitude,round_num),lat=round(Pickup_latitude,round_num)) %>% 
  count() %>% arrange(desc(n)) %>% head(5)


greentaxi <- makeIcon(
  iconUrl = "https://i.imgur.com/6rw618Q.png",
  iconWidth = 38, iconHeight = 35,
  iconAnchorX = 19, iconAnchorY = 39
)

Weekday_Top5 %>%
  leaflet() %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addCircleMarkers(~lng, ~lat, radius = 1,
                   color = "firebrick", fillOpacity = 0.001)%>%
  addMarkers(~lng, ~lat, icon = greentaxi)

Weekend_Top5 %>%
  leaflet() %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>%
  addCircleMarkers(~lng, ~lat, radius = 1,
                   color = "firebrick", fillOpacity = 0.001)%>%
  addMarkers(~lng, ~lat, icon = greentaxi)
```