---
title: "Gale Partner Interview Case Study"
author: "Chengran Ouyang"
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: lumen
    number_sections: true
    font-family: Open Sans, sans-serif
    font-import: https://fonts.googleapis.com/css?family=Open+Sans
    # code_folding: hide
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Define the Problem
**Requirment:**
1. Download the data, load it into your favorite statistical programing software or database.
Report the number of rows and columns that you've loaded.
2. Visualize trip distance by time of day in any way you see fit, any observations?
3. What are the most popular pickup locations on weekdays vs weekend?
4. I want to know where I can most easily get a cab. Recommend a pick up spot where i can
find a cab given my lat long.

Source: <http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml>.

The scope of data is limited to Green taxi data for February 2016. There is also data dictionary available for the website.

## Dependencies {.tabset}
### Required libraries
```{r }
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, DT, lubridate, leaflet, leaflet.extras, maps, data.table, ggthemes)
```

### Required Dataset

```{r }
# data <- read.csv("green_tripdata_2016-02.csv", stringsAsFactors = F)
# write_csv(data,"green_tripdata_2016-02v2.csv")
data <- fread("green_tripdata_2016-02v2.csv", stringsAsFactors = F, data.table = FALSE, na.strings=c("NA","NaN","?", ""),
              showProgress = F)
```


